Learning a functional distance between organizations with unsupervised embeddings of scientific mobility

### 2020-07-24 paper/current/main.tex
updated with table, new supp figure, and new text related to fig 6

Human mobility and migration drives major social phenomena such as the growth and evolution of cities~\autocite{wef2017migration, curiel2018citygravity},  epidemics~\autocite{kraemer2020covid, truscott2012epidemicgravity, xia2005measlesgravity}, economies~\autocite{kaluza2010cargo, kerr2011immigration}, innovation~\autocite{kaiser2018innovation, armano2017innovation},  and scientific research~\autocite{sugimoto2017mostimpact, petersen2018multiscale, franzoni2014advantage, morgan2018prestige, rodrigues2016mobility}. 
Common models of mobility, such as the Gravity Model~\autocite{zipf1946gravity, xia2005measlesgravity, jung2008highwaygravity, barbosa2018human, curiel2018citygravity, lewer2008immigrationgravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity} aim to capture mobility between places using their population and the \textit{distance} between them.
However, geographic distance is less relevant in the increasingly-globalized world~\autocite{czaika2018migration, czaika2018globalisation, buchan2009globalization, altbach2004globalization} wherein the physical barrier shrinks while linguistic, economic, cultural, and historical similarities matter more.
Alternative notions of functional distance~\autocite{brown1970migration, brown1970functional, kim2018functional} attempt to re-define distance by using some of aforementioned factors~\autocite{boschma2005proximity,  brown1970functional, brown1970migration, kim2018functional} but are often arbitrary and consider only single isolated facets. 
Here, using a rich dataset of scientific mobility, we demonstrate that a data-driven neural embedding approach~\autocite{mikolov2013word2vec} can successfully encodes a functional distance between organizations and captures the latent and multi-faceted structure of scientist's mobility.
We demonstrate that the cosine distance between organization embedding vectors, constructed using the affiliation trajectories of scientists in the Web of Science database, better explain and predict real mobility than geographic distance. 
Exploiting their unique properties, we demonstrate that the embedding space encodes not only geographic relations between organizations, but also relations based on culture, language, and academic prestige at multiple levels of granularity.
Also captured in the embedding space is a universal pattern capturing the relationship between organization's size and their connectedness.
The ability to learn scalable, dense, and meaningful high-resolution vector representations of mobility directly from the data can open up a new avenue of studying human mobility.

@@

How far apart are two places?
The question is surprisingly hard to answer when it involves human mobility.
While geographic distance may have driven human movements in the past, it holds less importance to a world connected by rapid transit, complex international relations, and dense airline networks.
For instance, a person living in Australia is more likely to migrate to the United Kingdom, a far-away country with similar language and culture, than to much closer countries like Indonesia. 
Similarly, a student in South Korea is more likely to attend universities in Canada than those in North Korea. 
Although geographic distance has still been used as the basis for the most prominent models of mobility, such as the Gravity~\autocite{zipf1946gravity} and Radiation~\autocite{simini2012universal} model,  \textit{functional distances}~\autocite{boschma2005proximity, brown1970functional, brown1970migration, kim2018functional} have been proposed to learn alternative notions of distance from real-world data or \textit{a priori} relationships between geographic entities. 
Yet, functional distances are often low-resolution and rely on somewhat arbitrary decisions about the importance of each factor.
Moreover, functional distances typically only focus on a single facet of mobility at a time, whereas real-world mobility is multi-faceted, influenced simultaneously by geography, language, culture, history, and economic opportunity. 
To capture such complexity of human mobility, new approaches are needed to incorporate its many facets into a single high-resolution functional distance.

@@

We use scientific mobility, inferred from affiliations, to demonstrate the effectiveness of our approach.
Scientific mobility is a central driver of the increasingly-globalized scientific enterprise~\autocite{czaika2018globalisation, altbach2004globalization, box2008competition}, being strongly related to innovation~\autocite{braunerhjelm2020labor, kaiser2018innovation, armano2017innovation}, impact~\autocite{sugimoto2017mostimpact, petersen2018multiscale, franzoni2014advantage}, collaboration~\autocite{rodrigues2016mobility}, and the diffusion of knowledge~\autocite{braunerhjelm2020labor, azoulay2011diffusion, morgan2018prestige}.
Moreover, scientific mobility has known structure, such as the centrality of the United States and the United Kingdom, and the strong prestige hierarchy~\autocite{auriol2010careers, clauset2015hierarchy, deville2014career} that should be captured in a representation.
However, understandings of scientific mobility have been limited by the sheer scope and complexity of the phenomenon~\autocite{robinson2019mobility, vannoorden2012mobility, deville2014career, ackers2008mobility, scott2015dynamics} and one of the underlying these challenges is the diminishing role of geography in shaping the landscape of scientific mobility.

@@

The embedding distance explains and predicts scientific mobility better than geographic distance.
The amount of variance explained by embedding distance ($R^{2} = 0.46$, Fig.~\ref{fig:gravity_performance}a) is more than twice that of geographic distance ($R^{2} = 0.22$, Fig.~\ref{fig:gravity_performance}b).
This pattern holds when considering the subset of only domestic (Fig.~\ref{fig:supp:gravity_correlation}c) and only international mobility flows (Fig.~\ref{fig:supp:gravity_correlation}d).
After fitting model parameters from the data, the embedding distance produces better predictions of expected flux than does geographic distance (Fig.~\ref{fig:gravity_performance}c--d), which also holds for the subsets of only domestic (Fig.~\ref{fig:supp:gravity_correlation}e--f) and international mobility (Fig.~\ref{fig:supp:gravity_correlation}g--h).
The embedding distance defined from the cosine distance between vectors also outperforms that defined from their dot product ($R^{2} = 0.34$, Fig.~\ref{fig:supp:gravity_dot}).
The embedding distance also out-performs alternative ``diffusion-based'' network distance measures such as the personalized-page rank scores (see supporting materials) calculated from the underlying mobility network (Fig.~\ref{fig:supp:gravity_pprcos}, Fig.~\ref{fig:supp:gravity_pprjsd}). A reasonable explanation behind these results is that raw network distance measures cannot capture high-order interaction because the network is the first-order Markov process model of the system, while the neural embedding can leverage it. Also, raw network distance measures are well-known for vulnerable on random error \autocite{holme2002attack}.
Our results unequivocally argue that the distance learned by the embedding method better captures actual mobility patterns than does geographic distance.

%
% Figure - Gravity law performance
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/GravityLawFig/gravity_performance.pdf}
	\caption{
		\textbf{Embedding distance outperforms geographic distance.}
		\textbf{a.} embedding distance (cosine distance between organization vectors) better explains the expected flux between organization pairs than geographic distance (\textbf{b}).
		The red line is the line of the best fit.
		Black dots are mean flux across binned distances.
		Error bars represent 99\% confidence intervals.
		The lightness of each hex bin indicates the frequency of organization pairs within it.
		\textbf{c-d.} Predictions of flux between organization pairs made using embedding distance outperform those made using geographic distance.
		Boxplots show distribution of actual flux for binned values of predicted flux.
		Box color corresponds to the degree to which the distribution overlaps $y = x$.
		``RMSE'' is root-mean-squared error between the actual and predicted values.
		\textbf{e-f.} Predictions made using model parameters learned from a subset of domestic (within-country) organization pairs.
		Embedding distance outperforms those made for geographic distance.
		\textbf{g-h.} Embedding distance also outperforms geographic distance when predicting flux using model parameters estimated from only international mobility (across-country organization pairs).
		Predictions are not meaningfully different if estimates are learned on all pairs of organizations (Fig. \ref{fig:supp:gravity_prediction_subsets}).
		These predictions are robust whether the model is fit on the global data, or on the domestic (Fig.~\ref{fig:gravity_performance}e-h) or international (Fig.~\ref{fig:supp:gravity_prediction_subsets}) subsets.
	}
	\label{fig:gravity_performance}
\end{figure}

@@

The embedding space also allows us to \textit{zoom in} on subsets and re-project them in order to reveal more local relationships. 
For example, re-projecting organizations located in Western, Southern, and Southeastern Asia (Fig.~\ref{fig:projection}b) reveals a gradient of countries between Egypt to Philippines that corresponds largely to geography, but with some peculiar exceptions based on cultural and religious spectrum.
For example, countries like Iraq are nearer to some southeast-Asian countries like Malaysia, with a cosine distance between the nation's aggregate vectors of $d = 0.29$, than to nearer and more religiously-similar nations like Saudi Arabia ($d = 0.50$) and Qatar ($d = 0.51$). 
Linguistic and historical ties are also conspicuous.
For example, we observe that Spanish-speaking post-colonial Latin American nations are positioned near Spain (Fig.~\ref{fig:projection}c), rather than nearby Portuguese-speaking nations ($d = 0.34$ vs. $d = 0.52$ for Mexico and $d = 0.39$ vs. $d = 0.49$ for Chile) reflecting linguistic and cultural constraints.
Similarly, North-African countries that were once under French rule such as Morocco are nearer to France ($d =0.35$) than to similarly-distant European countries as Spain ($d = 0.40$) and Italy ($d = 0.53$). 
Comparable structural patterns exist even within a single country.
For example, at the global level (Fig~\ref{fig:projection}a), organizations within Quebec in Canada are located nearer the French cluster ($d = 0.65$) than the rest of the Canadian universities ($d = 0.88$).

@@

What is it about the organization embedding that captures prestige so effectively?
We find that the size (L2 norm) of the organization embedding vectors provides insights into the characteristics of organizations. 
Up to a point, the size of U.S. organization's vectors tends to increase proportional to its number of researchers (both mobile and non-mobile) who published with that organization;
these organizations are primarily teaching-focused institutions, agencies, and hospitals that either are not ranked or have a low ranking. 
However, at around 1,000 researchers, the size of the vector decreases as the number of researchers increases;
organizations here are primarily research-intensive and prestigious universities with higher rankings, research outputs, R\&D funding, and doctoral students (Fig.~\ref{fig:supp:length_vs_metaInfo}).
A similar pattern has been observed in applications of neural embedding to natural language, wherein the size of word vectors were found to represent word ``significance''~\autocite{schakel2015measuring}.
If a word appears frequently in a specific context, the norm of the word vector is likely to be large, because a large norm is associated with how predictably the word occurs given a context.
However, if the word in question is universal, appearing frequently in many different contexts, it cannot have a large norm as it lacks a strong association with a particular context. 
Likewise, an organization such as Harvard with a small norm appears in many contexts, alongside many different organizations in affiliation trajectories. 
The concavity emerges in part from the relationship between the size of the vector and the expected connectedness of the organization given its size (Fig.~\ref{fig:length}a, inset, $R^{2} = 0.17$).
Large, prestigious, and well-funded research universities such as Princeton and Harvard have smaller vector sizes because they appear in many different contexts compared to more teaching-focused organizations such as NY Medical College, and University of Michigan at Flint. 
Some universities, such as the University of Alaska at Fairbanks, have surprisingly small vectors, perhaps as a result of its remote location and unique circumstances.

This same convex-curve repeats across many countries, though with variations that can reveal characteristics about the scientific systems of each country (Fig.~\ref{fig:length}b). 
For example, Chinese organizations have a longer average embedding vector size ($\bar{l} = 4.81$) than organizations in the United States ($\bar{l} = 0.421$), reflecting the tendency for organizations in the United States to appear more frequently in more different contexts than Chinese organizations. 
Smaller but scientifically advanced countries such as Sweden, and other large populous countries like Brazil exhibit curves similar to the United States (Fig.~\ref{fig:length}b, inset), a pattern repeated across most of the 20 countries with the most researchers (Fig.~\ref{fig:concave20}). 
Other nations exhibit different curves, such as Egypt which has an "S"-shaped curve---their largest organizations do not have smaller vectors, as in many other countries. 
Other nations such as Poland and Turkey demonstrate a linear relationship between an organization's number of researchers and the vector size, indicating that their largest organizations belong to very specific contexts (Fig.~\ref{fig:concave20}).
By encoding whether an organization will appear in specific or many different contexts, organization embeddings offer a way to characterize universal and particularistic patterns of mobility across countries.

%
% Figure - Boomerang
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Math/boomerang.pdf}
	\caption{
		\textbf{Size of organization embedding vectors captures connection diversity of organizations.}
		\textbf{a.} Size (L2 norm) of organization embedding vectors compared to their number of researchers for U.S. universities.
		Color indicates the rank of the university from the Times ranking, with 1 being the highest ranked university.
		Uncolored points are universities not listed on a Times ranking.
		A convex-shape emerges, wherein larger universities tend to more distant from the origin (large L2 norm), however the more prestigious the university, the closer it is to the origin (small L2 norm). 
		The L2 norm is correlated weakly with the ratio between organization's of number of connected organizations and its number of researchers ($R^{2} = 0.17$, \textbf{inset}).
		The red line is the line of the best fit, and 99\% confidence intervals are shown. 
		\textbf{b.} This convex pattern repeats globally for all organizations across all countries, shown for the United States, China, and others,
		The loess regression lines are shown for each selected country, and for the aggregate of remaining countries, with 99\% confidence intervals. 
		Loess lines are also shown for organizations in Sweden, Brazil, and Egypt (inset).
	}
	\label{fig:length}
\end{figure}

@@

\subsection*{Embedding}
%% Mobility sentences
We embed co-affiliation trajectories by treating them analogously to sentences and organizations analogously to words.
The trajectories of mobile authors were converted to ``affiliation trajectories'' by concatenating together their ordered list of unique organization identifiers, as demonstrated in Fig.~\ref{fig:methods:mobility_sentence}.a.
In more complex cases, such as listing multiple affiliations on the same paper or publishing with different affiliations on multiple affiliations in the same year, the order is randomized within that year, as shown in  Fig.~\ref{fig:methods:mobility_sentence}.b.

%% word2vec
These affiliation trajectories are used as input to the standard skip-gram negative sampling word embedding, commonly known as \textit{word2vec}~\autocite{mikolov2013word2vec} .
\textit{word2vec} constructs dense vector representations of words and phrases, in which distance relationships between words correspond to a notion of semantic distance. There are two types of embedding vectors in the model; in-vector and out-vector. In-vector is usually mainly used as a representation vector of each word \autocite{linzhuo2020hyperbolic, tshitoyan2019mat2vec, garg2018gender, kozlowski2018geometry, hamilton2016diachronic, le2014doc2vec, nakandala2016twitch} and we also used in-vector in this study.
By embedding affiliation trajectories, we aim to learn a dense vector for every organization where the distance between vectors relates to the tendency for two organizations to co-occur in a token window around the target organization token in the same career trajectory.
Vectors are learned by training a neural network to optimize the prediction of the so-called context term given a target, in our case using one organization to predict the others appearing in the same co-affiliation trajectory.
The hidden layer of the trained neural network is used as the final representation.

@@

\newpage
%
% Figure - publications over time
%
\begin{figure}[p!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Descriptive/pubs_over_time.pdf}
	\caption{
		\textbf{Publications over time.}
		\textbf{a.}
		The number of papers published by mobile authors has been steadily increasing from 2008 to 2017, with a small decrease in 2018,  likely an artifact of the Web of Science indexing process.
		Lines correspond to publications by mobile authors (org, purple), by authors with affiliations in at least two cities (blue), at least two regions (green), and at least two countries (yellow).
		There is no evidence of major changes in the publication patterns of mobile authors during this time period.
		\textbf{b.}
		Lines correspond to the proportion of publications classified as Biology and Health (black), Physics and Engineering (purple), Life and Earth Science (magenta), Social Science and Humanities (orange), and Math and Computer Science (yellow).
		The rate of publication in Biology and Health has leveled since about 2013, whereas the rate of publication in other fields has steadily increased.
		\textbf{c.}
		Despite increasing absolute counts, the percentage of mobile scholars, and those with at least two city, regional, or country affiliations, as a proportion of all publications, has decreased over time.
		\textbf{d.}
		The proportion of authors' publications across fields has largely remained steady.
		Biology and Health Science has comprised the majority of publications across nearly all years but has steadily declined in proportion.
		However, the proportion of Social Science and Humanities publications has been steadily increasing.
	}
	\label{fig:supp:pubs_over_time}
\end{figure}

@@

%
% Figure - Concave-curves countries
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Math/boomerang_20_countries.pdf}
	\caption{
		\textbf{Concave-curve repeats across most of 20 countries with most researchers.}
		Size (L2 norm) of organization embedding vectors compared to their number of researchers for U.S. universities.
		Loess regression line is shown for each country with 99\% confidence intervals. 
		Countries shown are the top 20 countries with the largest number of total unique researchers, as derived from our bibliographic data. 
	}
	\label{fig:concave20}
\end{figure}

>>>>

Human mobility and migration drives major social phenomena such as the growth and evolution of cities~\autocite{wef2017migration, curiel2018citygravity},  epidemics~\autocite{kraemer2020covid, truscott2012epidemicgravity, xia2005measlesgravity}, economies~\autocite{kaluza2010cargo, kerr2011immigration}, innovation~\autocite{kaiser2018innovation, armano2017innovation},  and scientific research~\autocite{sugimoto2017mostimpact, petersen2018multiscale, franzoni2014advantage, morgan2018prestige, rodrigues2016mobility}. 
Common models of mobility, such as the Gravity Model~\autocite{zipf1946gravity, xia2005measlesgravity, jung2008highwaygravity, barbosa2018human, curiel2018citygravity, lewer2008immigrationgravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity} aim to capture mobility between places using their population and the \textit{distance} between them.
However, geographic distance is less relevant in the increasingly-globalized world~\autocite{czaika2018migration, czaika2018globalisation, buchan2009globalization, altbach2004globalization} wherein the physical barrier shrinks while linguistic, economic, cultural, and historical similarities matter more.
Alternative notions of functional distance~\autocite{brown1970migration, brown1970functional, kim2018functional} attempt to re-define distance by using some of the aforementioned factor~\autocite{boschma2005proximity,  brown1970functional, brown1970migration, kim2018functional} but are often arbitrary, and consider only single isolated facets. 
Here, using a rich dataset of scientific mobility, we demonstrate that a data-driven neural embedding approach~\autocite{mikolov2013word2vec} can successfully encodes a functional distance between organizations and captures the latent and multi-faceted structure of scientist's mobility.
We demonstrate that the cosine distance between organization embedding vectors, constructed using the affiliation trajectories of scientists in the Web of Science database, better explain and predict real mobility than geographic distance. 
Exploiting their unique properties, we idemonstrate that the embedding space encodes not only geographic relations between organizations, but also relations based on culture, language, and academic prestige at multiple levels of granularity.
Also captured in the embedding space is a universal pattern capturing the relationship between organization's size and their connectedness.
The ability to learn scalable, dense, and meaningful high-resolution vector representations of mobility directly from the data can open up a new avenue of studying human mobility.

@@

How far apart are two places?
The question is surprisingly hard to answer when it involves human mobility.
While geographic distance may have driven human movements in the past, it holds less importance to a world connected by rapid transit, complex international relations, and dense airline networks.
For instance, a person living in Australia is more likely to migrate to the United Kingdom, a far-away country with similar language and culture, than to much closer countries like Indonesia. 
Similarly, a student in South Korea is more likely to attend universities in Canada than those in North Korea. 
Although geographic distance has still been used as the basis for the most prominent models of mobility, such as the Gravity~\autocite{zipf1946gravity} and Radiation~\autocite{simini2012universal} model,  \textit{functional distances}~\autocite{boschma2005proximity, brown1970functional, brown1970migration, kim2018functional} have been proposed to learn alternative notions of distance from real-world data or \textit{a priori} relationships between geographic entities. 
Yet, functional distances are often low-resolution and rely on somewhat arbitrary decisions about the importance of each factor.
Moreover, functional distances typically only focus on a single facet of mobility at a time, whereas real-world mobility is multi-faceted, influenced simultaneously by geography, language, culture, history, and economic opportunity. 
To capture the complexity of human mobility, and so new approaches are needed to incorporate its many facets into a single high-resolution measure of functional distance.

@@

We use scientific mobility, inferred from affiliations, to demonstrate the effectiveness of our approach.
Scientific mobility is a central driver of the increasingly-globalized scientific enterprise~\autocite{czaika2018globalisation, altbach2004globalization, box2008competition}, being strongly related to innovation~\autocite{braunerhjelm2020labor, kaiser2018innovation, armano2017innovation}, impact~\autocite{sugimoto2017mostimpact, petersen2018multiscale, franzoni2014advantage}, collaboration~\autocite{rodrigues2016mobility}, and the diffusion of knowledge~\autocite{braunerhjelm2020labor, azoulay2011diffusion, morgan2018prestige}.
Moreover, scientific mobility has known structure, such as the centrality of the United States and the United Kingdom, and the strong prestige hierarchy~\autocite{auriol2010careers, clauset2015hierarchy, deville2014career} that should be captured in a representation.
However, understandings of scientific mobility have been limited by the sheer scope and complexity of the phenomenon~\autocite{robinson2019mobility, vannoorden2012mobility, deville2014career, ackers2008mobility, scott2015dynamics}.
Underlying these challenges is the diminishing role of geography and geographic distance in shaping the landscape of scientific mobility.

@@

The embedding distance explains and predicts scientific mobility better than geographic distance.
The amount of variance explained by embedding distance ($R^{2} = 0.46$, Fig.~\ref{fig:gravity_performance}a) is more than twice that of geographic distance ($R^{2} = 0.22$, Fig.~\ref{fig:gravity_performance}b).
This pattern holds when considering the subset of only domestic (Fig.~\ref{fig:supp:gravity_correlation}c) and only international mobility flows (Fig.~\ref{fig:supp:gravity_correlation}d).
After fitting model parameters from the data, the embedding distance produces better predictions of expected flux than does geographic distance (Fig.~\ref{fig:gravity_performance}c--d), which also holds for the subsets of only domestic (Fig.~\ref{fig:supp:gravity_correlation}e--f) and international mobility (Fig.~\ref{fig:supp:gravity_correlation}g--h).
The embedding distance defined from the cosine distance between vectors also outperforms that defined from their dot product ($R^{2} = 0.34$, Fig.~\ref{fig:supp:gravity_dot}).
The embedding distance also out-performs alternative ``diffusion-based'' network distance measures such as the personalized-page rank scores (see supporting materials) calculated from the underlying mobility network (Fig.~\ref{fig:supp:gravity_pprcos}, Fig.~\ref{fig:supp:gravity_pprjsd}).
Our results unequivocally argue that the distance learned by the embedding method better captures actual mobility patterns than does geographic distance.

%
% Figure - Gravity law performance
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/GravityLawFig/gravity_performance.pdf}
	\caption{
		\textbf{Embedding distance outperforms geographic distance.}
		\textbf{a.} embedding distance (cosine distance between organization vectors) better explains the expected flux between organization pairs than geographic distance (\textbf{b}).
		The red line is the line of the best fit.
		Black dots are mean flux across binned distances.
		Error bars are 99\% confidence intervals.
		The lightness of each hex bin indicates the frequency of organization pairs within it.
		\textbf{c-d.} Predictions of flux between organization pairs made using embedding distance outperform those made using geographic distance.
		Boxplots show distribution of actual flux for binned values of predicted flux.
		Box color corresponds to the degree to which the distribution overlaps $y = x$.
		``RMSE'' is root-mean-squared error between the actual and predicted values.
		\textbf{e-f.} Predictions made using model parameters learned from a subset of domestic (within-country) organization pairs.
		Embedding distance outperforms those made for geographic distance.
		\textbf{g-h.} Embedding distance also outperforms geographic distance when predicting flux using model parameters estimated from only international mobility (across-country organization pairs).
		Predictions are not meaningfully different if estimates are learned on all pairs of organizations (Fig. \ref{fig:supp:gravity_prediction_subsets}).
		These predictions are robust whether the model is fit on the global data, or on the domestic (Fig.~\ref{fig:gravity_performance}e-h) or international (Fig.~\ref{fig:supp:gravity_prediction_subsets}) subsets.
	}
	\label{fig:gravity_performance}
\end{figure}

@@

The embedding space also allows us to \textit{zoom in} on subsets and re-project them in order to reveal more local relationships. 
For example, re-projecting organizations located in Western, Southern, and Southeastern Asia (Fig.~\ref{fig:projection}b) reveals a gradient of countries between Egypt to Philippines that corresponds largely to geography, but with some peculiar exceptions. 
For example, countries like Iraq are nearer to some southeast-Asian countries like Malaysia, with a cosine distance between the nation's aggregate vectors of $d = 0.29$, than to nearer and more religiously-similar nations like Saudi Arabia ($d = 0.50$) and Qatar ($d = 0.51$). 
Linguistic and historical ties are also conspicuous.
For example, we observe that Spanish-speaking post-colonial Latin American nations are positioned near Spain (Fig.~\ref{fig:projection}c), rather than nearby Portuguese-speaking nation ($d = 0.34$ vs. $d = 0.52$ for Mexico and $d = 0.39$ vs. $d = 0.49$ for Chile) reflecting linguistic and cultural constraints.
Similarly, North-African countries that were once under French rule such as Morocco are nearer to France ($d =0.35$) than to similarly-distant European countries as Spain ($d = 0.40$) and Italy ($d = 0.53$). 
Comparable structural patterns exist even within a single country.
For example, at the global level (Fig~\ref{fig:projection}a), organizations within Quebec in Canada are located nearer the French cluster ($d = 0.65$) than the rest of the Canadian universities ($d = 0.88$).

@@

What is it about the organization embedding that captures prestige so effectively?
We find that the size (L2 norm) of the organization embedding vectors provides insights into the characteristics of organizations. 
Up to a point, the size of U.S. organization's vectors tends to increase proportional to its number of researchers (both mobile and non-mobile) who published with that organization;
these organizations are primarily teaching-focused institutions, agencies, and hospitals that either are not ranked or have a low ranking. 
However, at around 1,000 researchers, the size of the vector decreases as the number of researchers increases;
organizations here are primarily research-intensive and prestigious universities with higher rankings, research outputs, R\&D funding, and doctoral students (Fig.~\ref{fig:supp:length_vs_metaInfo}).
A similar pattern has been observed in applications of neural embedding to natural language, wherein the size of word vectors were found to represent word ``significance''~\autocite{schakel2015measuring}.
If a word appears frequently in a specific context, the norm of the word vector is likely to be large, because a large norm is associated with how predictably the word occurs given a context.
However, if the word in question is universal, appearing frequently in many different contexts, it cannot have a large norm as it lacks a strong association with a particular context. 
Likewise, an organization such as Harvard with a small norm appears in many contexts, alongside many different organizations in affiliation trajectories---it is well-connected. 
The concavity emerges in part from the relationship between the size of the vector and the expected connectedness of the organization given its size (Fig.~\ref{fig:length}a, inset, $R^{2} = 0.17$).
Large, prestigious, and well-funded research universities such as Princeton and Harvard have smaller vector sizes because they appear in many different contexts compared to more teaching-focused organizations such as NY Medical College, and University of Michigan at Flint. 
Some universities, such as the University of Alaska at Fairbanks, have surprisingly small vectors, perhaps as a result of its remote location and unique circumstances.

This same convex-curve repeats across many countries, though with variations that can reveal characteristics about the scientific systems of each country (Fig.~\ref{fig:length}b). 
For example, the representative vector of Chinese organizations are larger than that of the United States ($\bar{l} = 2.97$ vs $\bar{l} = 2.39$, Table~\ref{table:supp:norm}), reflecting the tendency for organizations in the United States to appear more frequently in more different contexts than Chinese organizations. 
Smaller but scientifically advanced countries such as Sweden, and other large populous countries like Brazil exhibit curves similar to the United States (Fig.~\ref{fig:length}b, inset), a pattern repeated across most countries with large scientific systems (Fig.~\ref{fig:concave30}). 
Other nations exhibit different curves, such as Egypt which has an "S"-shaped curve---their largest organizations do not as well connected as other country's largest organizations. 
Other nations such as Poland, Iran, and Turkey demonstrate a linear relationship between an organization's number of researchers and the vector size, indicating that their largest organizations belong to very specific contexts (Fig.~\ref{fig:concave30}).
The organization-level distribution of embedding vector norms reveals even deeper heterogeneity---
for example, whereas the United States has the smallest representative vector, skewness of the sizes of its organization vectors suggest that this is driven by a small number of especially-well-connected organizations ($s = -0.82$, Fig.~\ref{fig:norm_dist}). 
Germany and the United Kingdom have comparable representative vector sizes to the United States ($\bar{l} = 2.6$ and $\bar{l} = 2.61$, respectively), though they are much less skewed ($s = -0.63$ and $s = -0.55$), suggesting less inequality in the connectedness of organizations in these countries. 
The vector sizes of organizations in some countries are even more skewed, such as in Iran ($\bar{l} = 3.57$, $s = -2.13$), and China ($\bar{l} = 2.97$, $s = -1.08$), indicating the strong difference between their most- and least-connected organizations.
For some countries, their organizations are positively-skewed, though seemingly for different reasons---
Austria, for example, has a balanced distribution of organization vector sizes, suggesting a diverse range of organizations with most being well connected ($\bar{l} =2.64$, $s = 0.18$);
Russia, in contrast, has a high number of organization vectors of moderate size, but also several very isolated organizations with large vector sizes ($\bar{l} = 3.08$, $s = 0.67$). 
By capturing the connectedness of organizations in the embedding vector size, the embedding space can be used to characterize universal and particularistic patterns of mobility across countries.

%
% Figure - Boomerang
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Math/boomerang.pdf}
	\caption{
		\textbf{Size of organization embedding vectors captures connection diversity of organizations.}
		\textbf{a.} Size (L2 norm) of organization embedding vectors compared to their number of researchers for U.S. universities.
		Color indicates the rank of the university from the Times ranking, with 1 being the highest ranked university.
		Uncolored points are universities not listed on a Times ranking.
		A convex-shape emerges, wherein larger universities tend to more distant from the origin (large L2 norm), however the more prestigious the university, the closer it is to the origin (small L2 norm). 
		The L2 norm is correlated weakly with the ratio between organization's of number of connected organizations and its number of researchers ($R^{2} = 0.17$, \textbf{inset}).
		The red line is the line of the best fit, and 99\% confidence intervals are shown. 
		\textbf{b.} This convex pattern repeats globally for all organizations across all countries, shown for the United States, China, and others,
		The loess regression lines are shown for each selected country, and for the aggregate of remaining countries, with 99\% confidence intervals. 
		Loess lines are also shown for organizations in Sweden, Brazil, and Egypt (\textbf{inset}).
	}
	\label{fig:length}
\end{figure}

@@

\subsection*{Embedding}
%% Mobility sentences
We embed co-affiliation trajectories by treating them analogously to sentences, and organizations analogously to words.
The trajectories of mobile authors were converted to ``affiliation trajectories'' by concatenating together their ordered list of unique organization identifiers, as demonstrated in Fig.~\ref{fig:methods:mobility_sentence}.a.
In more complex cases, such as listing multiple affiliations on the same paper or publishing with different affiliations on multiple affiliations in the same year, the order is randomized within that year, as shown in  Fig.~\ref{fig:methods:mobility_sentence}.b.

%% word2vec
These affiliation trajectories are used as input to the standard skip-gram negative sampling model, for word embedding, commonly known as \textit{word2vec}~\autocite{mikolov2013word2vec} .
In natural-language processing, word embeddings are used to construct dense vector-space representations of words and phrases, in which distance relationships between words correspond to a notion of semantic distance.
By embedding affiliation trajectories, we aim to learn a dense vector for every organization where the distance between vectors relates to the tendency for two organizations to co-occur in a token window around the target organization token in the same career trajectory.
Vectors are learned by training a neural network to optimize the prediction of the so-called context term given a target, in our case using one organization to predict the others appearing in the same co-affiliation trajectory.
The hidden layer of the trained neural network is used as the final representation.

@@

\begin{table}[]
\centering
\caption{\textbf{L2 Norm of country's representative vectors}.
			 Shown for top 30 countries with the most unique mobile and non-mobile researchers}
\label{table:supp:norm}
\begin{tabular}{lrr}
\textbf{Country} & \textbf{L2 Norm} & \textbf{\# Organizations} \\ \hline
United States    & 2.39             & 1281                      \\
Germany          & 2.6              & 485                       \\
United Kingdom   & 2.61             & 514                       \\
Austria          & 2.64             & 74                        \\
France           & 2.83             & 688                       \\
Belgium          & 2.84             & 84                        \\
Switzerland      & 2.85             & 66                        \\
Spain            & 2.94             & 322                       \\
China            & 2.97             & 497                       \\
India            & 2.99             & 114                       \\
Poland           & 3.02             & 145                       \\
Canada           & 3.02             & 147                       \\
Italy            & 3.04             & 386                       \\
Russia           & 3.08             & 187                       \\
Norway           & 3.1              & 122                       \\
Netherlands      & 3.11             & 136                       \\
Sweden           & 3.16             & 75                        \\
Brazil           & 3.16             & 286                       \\
Finland          & 3.17             & 66                        \\
Denmark          & 3.21             & 54                        \\
Czech Republic   & 3.23             & 97                        \\
Greece           & 3.24             & 62                        \\
Australia        & 3.24             & 90                        \\
Turkey           & 3.28             & 99                        \\
South Korea      & 3.28             & 156                       \\
Israel           & 3.32             & 71                        \\
Portugal         & 3.33             & 57                        \\
Japan            & 3.35             & 465                       \\
Iran             & 3.57             & 68                        \\
Taiwan           & 3.67             & 72                       
\end{tabular}
\end{table}

%
% Figure - publications over time
%
\begin{figure}[p!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Descriptive/pubs_over_time.pdf}
	\caption{
		\textbf{Publications over time.}
		\textbf{a.}
		The number of papers published by mobile authors has been steadily increasing from 2008 to 2017, with a small decrease in 2018,  likely an artifact of the Web of Science indexing process.
		Lines correspond to publications by mobile authors (org, purple), by authors with affiliations in at least two cities (blue), at least two regions (green), and at least two countries (yellow).
		There is no evidence of major changes in the publication patterns of mobile authors during this time period.
		\textbf{b.}
		Lines correspond to the proportion of publications classified as Biology and Health (black), Physics and Engineering (purple), Life and Earth Science (magenta), Social Science and Humanities (orange), and Math and Computer Science (yellow).
		The rate of publication in Biology and Health has leveled since about 2013, whereas the rate of publication in other fields has steadily increased.
		\textbf{c.}
		Despite increasing absolute counts, the percentage of mobile scholars, and those with at least two city, regional, or country affiliations, as a proportion of all publications, has decreased over time.
		\textbf{d.}
		The proportion of authors' publications across fields has largely remained steady.
		Biology and Health Science has comprised the majority of publications across nearly all years but has steadily declined in proportion.
		However, the proportion of Social Science and Humanities publications has been steadily increasing.
	}
	\label{fig:supp:pubs_over_time}
\end{figure}

@@

%
% Figure - Concave-curves countries
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Math/boomerang_30_countries.pdf}
	\caption{
		\textbf{Concave-curve repeats across most of 30 countries with most researchers.}
		Size (L2 norm) of organization embedding vectors compared to their number of researchers for U.S. universities.
		Loess regression line is shown for each country with 99\% confidence intervals. 
		Countries shown are the 30 with the largest number of total unique mobile and non-mobile researchers.
	}
	\label{fig:concave30}
\end{figure}
